### 2023-11-29 main.tex
abstract and cup

\begin{abstract}
\noindent
We study the boundary-driven version of the stirring process with $N-1$ species
on a generic graph $G=(V,\mathcal{E})$ with reservoirs. 
The multispecies stirring process generalizes the symmetric exclusion process, 
which is recovered in the case $N=2$.
We prove the existence of a dual process defined on
an extended graph $\widetilde{G}=(\widetilde{V},\widetilde{\mathcal{E})}$ which 
includes additional extra-sites $\widetilde{V}\setminus V$ where
dual particles gets absorbed in the long-time limit.  
We thus obtain a characterization of the non-equilibrium
steady state of the boundary-driven system in terms of the absorption
probabilities of dual particles.
For the case of one-dimensional chains with two reservoirs
at the boundaries and with maximum one particle per site,
we compute these absorption probabilities by relying
on the underlying ${gl}(N)$ symmetry and the
matrix product ansatz. Thus one gets a closed-formula
for (long-ranged) correlations and for the 
non-equilibrium stationary measure.
Extensions beyond this integrable set-up
are also discussed.
\end{abstract}

@@

Our strategy to solve for the non-equilibrium steady state is of a probabilistic nature and relies on the use of {\em duality} \cite{schutzSandow}. This approach, starting from the pioneering work of Kipnis, Marchiorio and Presutti \cite{KMP}, 
has been substantially developed in recent years to study stochastic processes in the boundary-driven set-up \cite{giardina2009duality,carinci2013duality,frassek2020duality,carinci2023open}. 
The boundary-driven process is mapped to a dual process with absorbing extra sites and the problem of describing the stationary state of the original system
is reduced to the problem of computing the absorption probabilities of the dual particles. 
In general, this is a difficult task. However,  for the case of integrable systems they can be computed exactly in closed-form. 
For recent examples in the framework of singlespecies systems, see \cite{frassek2020eigenstates, frassek2022exact}.  
Here we generalize the duality approach to the set-up of multispecies particle systems.

@@

{\bf Acknowledgments.} We acknowledge support from Istituto Nazionale di Alta Matematica (GNFM), Prin-project {\color{blue} xxxx}, FAR-project {\color{blue} xxxx} UniMoRe.

INFN Progetto Giovani

>>>>

\begin{abstract}
\noindent
We study the boundary-driven version of the stirring process with $N-1$ species
on a generic graph $G=(V,\mathcal{E})$ with reservoirs. 
The multispecies stirring process generalizes the symmetric exclusion process, 
which is recovered in the case $N=2$.
We prove the existence of a dual process defined on
an extended graph $\widetilde{G}=(\widetilde{V},\widetilde{\mathcal{E})}$ which 
includes additional extra-sites $\widetilde{V}\setminus V$ where
dual particles gets absorbed in the long-time limit.  
We thus obtain a characterization of the non-equilibrium
steady state of the boundary-driven system in terms of the absorption
probabilities of dual particles.
For the case of one-dimensional chains with two reservoirs
at the boundaries and with maximum one particle per site the process is integrable. Based on the results of \cite{vanicat2017exact},
we compute the absorption probabilities by relying
on the underlying ${gl}(N)$ symmetry and the
matrix product ansatz. Thus one gets a closed-formula
for (long-ranged) correlations and for the 
non-equilibrium stationary measure.
Extensions beyond this integrable set-up
are also discussed.
\end{abstract}

@@

Our strategy to solve for the non-equilibrium steady state is of a probabilistic nature and relies on the use of {\em duality} \cite{schutzSandow}. This approach, starting from the pioneering work of Kipnis, Marchiorio and Presutti \cite{KMP}, 
has been substantially developed in recent years to study stochastic processes in the boundary-driven set-up \cite{giardina2009duality,carinci2013duality,frassek2020duality,carinci2023open}. 
The boundary-driven process is mapped to a dual process with absorbing extra sites and the problem of describing the stationary state of the original system
is reduced to the problem of computing the absorption probabilities of the dual particles. 
In general, this is a difficult task. However,  for the case of integrable systems they can be computed exactly in closed-form. 
For recent examples in the framework of singlespecies systems, see \cite{frassek2020eigenstates, frassek2022exact}.  
{\color{red}
Here we generalize the duality approach to the set-up of multispecies particle systems.}

@@

{\bf Acknowledgments.} We acknowledge support from Istituto Nazionale di Alta Matematica (GNFM),  the INFN grant Gauge and String Theory (GAST),  the PRIN project CUP-E53D23002220006,  the FAR UNIMORE project CUP-E93C23002040005 and  the INdAM–GNFM project CUP-E53C22001930001.

% Prin-project {\color{blue} xxxx}, FAR-project {\color{blue} xxxx} UniMoRe.
% 
% INFN Progetto Giovani
